Brief Announcement: An Improved Lower Bound for Dynamic Balanced Graph Partitioning

### 2020-02-16 paper/paper.tex
Pesky paragraph rewritten!

  \medskip
  \maciek{The weakest paragraph}
  \maciek{x must be a sequence of requests, not the number; better name for x}
  Before we bound the competitive ratio, we relate the costs of $\TAlg$ and $\OPT$ with respect to requests $\sigma_E$.

  In a fixed configuration of $\OPT$, it might mitigate paying for requests between at most $3\ell$ pairs of nodes by collocating them in its clusters.
  Recall that $\sigma_E$ consist of requests between unsaturated pairs, and it accounts only for requests that increased the counter (i.e., external requests), thus $\OPT$ might mitigate at most $3\ell\cdot(\alpha - 1)$ requests from $\sigma_E$.
  Faced with

  We begin by lower-bounding the number of requests from $\sigma_E$ that would incur a cost for $\OPT$ if it .

  may  account (by either serving remotely or migrating), and we denote their number by $x$.
  It holds that $x \geq \TAlg(\sigma_E) - 3\cdot\ell\cdot(\alpha - 1)$, as it received all the requests that \TAlg did, but it could ignore the cost of serving at most $3\cdot\ell$ pairs of nodes, as it might have them collocated in its clusters.
  Each such pair of nodes account for at most $\alpha-1$ requests (the pairs from $\sigma_E$ are unsaturated) in the cost of $\TAlg(\sigma_E)$, hence collocating them would decrease their number by at most $3\cdot\ell\cdot(\alpha - 1)$.

  Now we bound the cost incurred by $\OPT$ while serving requests from $x$.
  Instead of serving all of them remotely, $\OPT$ can perform migrations to decrease its cost.
  By performing a swap of nodes $(u,v)$, $\OPT$ collocates $u$ with two nodes $u', u''$, and $v$ with two nodes $v'$, $v''$.
  This allows to serve requests between $(u,u')$, $(u,u'')$, $(v,v')$ and $(v,v'')$ for free afterwards.
  (express that between a pair of nodes we accounted $\alpha-1$ requests in $x$)
  Hence, by performing a migration for the cost $\alpha$, $\OPT$ does not pay the remote serving costs for at most $4\cdot (\alpha - 1)$ requests from $x$ ($x$ is not a sequence but its length!).
  Finally, for serving $x$, $\OPT$ pays at least $x \cdot \frac{\alpha}{4\cdot (\alpha-1)}$.
  This gives us
  \begin{equation*}
    \OPT(\sigma_E) + \OPT(M)  \geq (\TAlg(\sigma_E)-3\cdot\ell\cdot(\alpha - 1)) \cdot \frac{\alpha}{4\cdot (\alpha-1)}
    \geq \TAlg(\sigma_E) / 4 - 4\cdot \ell \cdot \alpha.
  \end{equation*}
  From that, we obtain $\TAlg(\sigma_E) \geq 4\cdot(\OPT(\sigma_E)+\OPT(M)) + 16\ell \cdot \alpha$.
  With that, we are ready to bound the competitive ratio.
  Let $E := \OPT(\sigma_E) + \OPT(M)$. Finally, using above bounds,
  \begin{equation*}
    \frac{\TAlg(\sigma)}{\OPT(\sigma)} \leq \frac{10\alpha \cdot \ell + \TAlg(\sigma_E)}{(\alpha + \OPT(\sigma_E) + \OPT(M))/2} \leq \frac{52\alpha\cdot\ell + 8\cdot E}{\alpha + E} \leq 52 \ell = O(\ell).
  \end{equation*}

>>>>

  \medskip
  Before we bound the competitive ratio, we relate the costs of $\TAlg$ and $\OPT$ with respect to requests $\sigma_E$.
  In a fixed configuration of $\OPT$, it might mitigate paying for requests between at most $3\ell$ pairs of nodes by collocating them in its clusters.
  Recall that $\sigma_E$ consists of requests between unsaturated pairs, and it accounts only for requests that increased the counter (i.e., external requests), thus $\OPT$ might mitigate at most $3\ell\cdot(\alpha - 1)$ requests from $\sigma_E$.
  Faced with $W := |\sigma_E| - 3\ell\cdot(\alpha-1)$ requests $\sigma_E$ that it could not mitigate, $\OPT$ served some of them remotely and possibly performed some migrations to decrease its cost.

  Now we estimate the cost of $\OPT(\sigma_E)$ by accounting savings from migrations.
  By performing a swap of nodes $(u,v)$, $\OPT$ collocates $u$ with two nodes $u', u''$, and $v$ with two nodes $v'$, $v''$.
  This allows to serve requests between $(u,u')$, $(u,u'')$, $(v,v')$ and $(v,v'')$ for free afterwards.
  As $\sigma_E$ consists of requests between unsaturated pairs, and it accounts only for external requests, there are at most $\alpha-1$ requests between each of these pairs.
  Thus, by performing a migration for the cost $\alpha$, $\OPT$ does not pay the remote serving costs for at most $4\cdot (\alpha - 1)$ requests from $\sigma_E$.
  Finally, for serving $\sigma_E$, $\OPT$ pays at least $W \cdot \frac{\alpha}{4\cdot (\alpha-1)}$.
  This gives us
  \begin{equation*}
    \OPT(\sigma_E) + \OPT(M)  \geq (\TAlg(\sigma_E)-3\cdot\ell\cdot(\alpha - 1)) \cdot \frac{\alpha}{4\cdot (\alpha-1)}
    \geq \TAlg(\sigma_E) / 4 - 4\cdot \ell \cdot \alpha.
  \end{equation*}
  From that, we obtain $\TAlg(\sigma_E) \geq 4\cdot(\OPT(\sigma_E)+\OPT(M)) + 16\ell \cdot \alpha$.
  With that, we are ready to bound the competitive ratio.
  Let $E := \OPT(\sigma_E) + \OPT(M)$. Finally, using above bounds,
  \begin{equation*}
    \frac{\TAlg(\sigma)}{\OPT(\sigma)} \leq \frac{10\alpha \cdot \ell + \TAlg(\sigma_E)}{(\alpha + \OPT(\sigma_E) + \OPT(M))/2} \leq \frac{52\alpha\cdot\ell + 8\cdot E}{\alpha + E} \leq 52 \ell = O(\ell).
  \end{equation*}
